---
title: "An Introduction to listr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An Introduction to listr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(listr)
```

The idea behind the `listr` package is to make working with lists with in R 
a bit more convenient. 
There is nothing wrong with lists in R per se, but sometimes the tools 
around lists and the syntax can be a bit confusing and complicated. 
With the tools provided by `listr` common tasks with lists and especially 
lists containing data frames can hopefully be simplified by providing a 
consistent and easy to read syntax that is also suited for use with pipes. 

For our examples below, assume we split the `mtcars` 
data by the `cyl` variable. 

```{r}
by_cyl <- split(mtcars, mtcars$cyl)
```


## Basic operations

```{r}
by_cyl <- by_cyl |> 
  list_rename("cyl4" = `4`, "cyl6" = `6`, "cyl8" = `8`)
```


```{r}
by_cyl |> list_select(cyl6)
```

There is both `list_extract` and `list_select` that can appear a bit similar 
in some cases. 
However, there are some important distinctions between both:

* `list_select` allows you to select several items from a list.
* `list_select` always returns a list.

In contrast

* `list_extract` only allows you to select a single item.
* `list_extract` unlists the selected item.

Thus, `list_select` is equivalent to selecting from a list with a single 
square bracket, while `list_extract` is equivalent to using double square 
brackets. 

```{r}
by_cyl |> list_select(1, 2)
```

```{r}
cyl4 <- by_cyl |> list_extract(cyl4)
cyl4
```

The `list_remove` function is straightforward.

```{r}
by_cyl <- by_cyl |> list_remove(cyl4)
by_cyl
```

The opposite can be achieved with `list_insert` and its more specialised
versions `list_append` and `list_prepend`. 

```{r}
by_cyl <- by_cyl |> list_prepend(cyl4, name = "cyl4")
by_cyl
```


## Operations with data frames
There is a certain focus on operations with data frames in this package. 

The function `list_name_to_df` adds a column to each data frame in the list
containing the name of the list item.
This is particularly useful if you have a list where each item is data from 
an experimental group or something similar.

```{r}
by_cyl |> 
  list_name_to_df() |> 
  list_select(1)
```

Using `list_bind` you can bind together data frame elements of a list. 
This is roughly similar to calling `do.call(rbind, list)` or the same with 
`cbind`, but there is a bit more flexibility. 

```{r}
by_cyl |> 
  list_bind(cyl4, cyl6, what = "rows", name = "cyl4_and_6")
```

As the `list_bind` function wraps calls to `rbind` and `cbind` it will
have the same effect on names as calling those two functions directly. 

Finally, there is also `list_join_df` allowing to form a single element 
out of many by joining data on an index.


## Flattening
Finally, there is `list_flatten` which takes in a list with nested list and
flattens it. 
The depth of the flattening process can be specified, by default all items
are moved to one level.

```{r}
foo <- list(
  1,
  list(1, 2, 3, 
       list(4, 5, 6),
       list(7, 8, 9, 
            list(10, 11, 
                 list(12))),
       13, 14),
  15
)

foo |> list_flatten()
```

```{r}
foo |> list_flatten(max_depth = 1)
```


